Compositional Equivalences Based on Open pNets

### 2022-11-23 2016-Weak-FH-Bisim/WeakBisim.tex
Other corrections

We let $e_i$ range over expressions ($e_i\in\AlgE$), $a$
range over action labels, $\symb{op}$ be operators, and $x_i$ and $y_i$ range over
variable names. We additionally rely on a set of action names, ranged over by $a$, $b$, \ldots

We define two kinds of parameterised actions. The first kind supports two kinds of parameters: input parameters that are variables and output parameters that can be any expression. The second kind makes no distinction between input and output parameters. 
The  actions that distinguish input variables will be used in the definition of \pLTS\ below and are defined as follows:
\[
\begin{array}[l]{rcl@{\quad}p{5.65cm}}
  \alpha\in\AlgA&::=&a(p_1,\ldots,p_n)&\text{action terms}\\
  p_i&::=& ?x~|~e_i&\text{parameters (input var or expression)}\\
  e_i&::=& \symb{Value}~|~x~|~\symb{op}(\symb{e}_1,..,\symb{e}_n)&\text{Expressions}
\end{array}
\]
The \emph{input variables} in an action term are those marked with a
$\symb{?}$.
We additionally impose that each input variable does not
appear anywhere else in the same action term:
$p_i=?x\Rightarrow\forall j\neq i.\, x\notin \vars(p_j)$.
We define $iv(t)$  as the set of input variables of a term $t$ (without the '?' marker).
Input variables are used in guards and to update the local state, they can only appear in well-identified expressions.
Action algebras can encode naturally usual point-to-point message passing calculi (using 
$a(?x_1,...,?x_n)$ for inputs, $a(v_1,..,v_n)$ for outputs), but they also allow
for more general synchronisation mechanisms, like gate negotiation in Lotos, or broadcast
communications.

@@

\begin{example}
The simulation of one transition by many others is one non-standard aspect of this definition. This is made necessary by the parameterised nature of our model. Consider the following open transition.
\[ \openrule
		{  \set{\beta}, \True,\subst{y\leftarrow x}}
          {s_1 \OTarrow {\alpha(x)} s_1'}\]
Bisimulation should allow it to be matched by the two following ones (depending on the value of $x$), to prove that the relation $\mathcal{R}=\{(s_1,s_2,\True),(s_1',s_2',\True)\}$ is a bisimulation.
\[ \openrule
         {
           \set{\beta}, x\geq 0,\subst{y\leftarrow x}}
          {s_2 \OTarrow {\alpha(x)} s_2'}
\qquad
\openrule
         {
           \set{\beta}, x<0,\subst{y\leftarrow x}}
          {s_2 \OTarrow {\alpha(x)} s_2'}
\]
This example illustrates the necessity of multiple transitions in the definition of bisimulation in a naive and minimalistic way.  It can easily be extended into a non-trivial example with more states and different usage of the variables.
\end{example}

@@

 $ \forall n\ge 1.  WI_{8b}(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(??\_msg,s\_ec}+n)\}, True, ()}
         {220 \OTWeakarrow {\nounderline{\texttt{out(??\_msg,m\_ec}+n)}} \{202, 000\}}
         $

>>>>

We let $e_i$ range over expressions ($e_i\in\AlgE$), $\symb{op}$ be operators, and $x_i$ and $y_i$ range over
variable names.  We additionally rely on a set of action names, ranged over by $a$, $b$, \ldots. 
We define two kinds of parameterised actions. The first kind supports two kinds of parameters: input parameters that are variables and output parameters that can be any expression. The second kind makes no distinction between input and output parameters. 
The  actions that distinguish input variables will be used in the definition of \pLTS\  and are defined as follows:
\[
\begin{array}[l]{rcl@{\quad}p{5.65cm}}
  \alpha\in\AlgA&::=&a(p_1,\ldots,p_n)&\text{action terms}\\
  p_i&::=& ?x~|~e_i&\text{parameters (input var or expression)}\\
  e_i&::=& \symb{Value}~|~x~|~\symb{op}(\symb{e}_1,..,\symb{e}_n)&\text{Expressions}
\end{array}
\]
The \emph{input variables} in an action term are those marked with a
$\symb{?}$.
We additionally impose that each input variable does not
appear anywhere else in the same action term:
$p_i=?x\Rightarrow\forall j\neq i.\, x\notin \vars(p_j)$.
We define $iv(t)$  as the set of input variables of a term $t$ (without the '?' marker).
Input variables are used in guards and to update the local state, they can only appear in well-identified expressions.
Action algebras can encode naturally usual point-to-point message passing calculi (using 
$a(?x_1,...,?x_n)$ for inputs, $a(v_1,..,v_n)$ for outputs), but they also allow
for more general synchronisation mechanisms, like gate negotiation in Lotos, or broadcast
communications.

@@

Note that, if there is a FH-bisimulation $\mathcal{R}$ such  that $(s,t|\Pred)\in\mathcal{R}$, and additionally $\Pred\,'\Rightarrow \Pred$ and  it does not mean that there is a FH-bisimulation $\mathcal{R}'$ such that $(s,t|\Pred\,')$, indeed pair of states such that $\Pred$ is true and $\Pred\,'$ is false might be necessary to prove the FH-bisimulation.
\begin{example}
The simulation of one transition by many others is one non-standard aspect of this definition. This is made necessary by the parameterised nature of our model. Consider the following open transition.
\[ \openrule
		{  \set{\beta}, \True,\subst{y\leftarrow x}}
          {s_1 \OTarrow {\alpha(x)} s_1'}\]
Bisimulation should allow it to be matched by the two following ones (depending on the value of $x$), to prove that the relation $\mathcal{R}=\{(s_1,s_2,\True),(s_1',s_2',\True)\}$ is a bisimulation.
\[ \openrule
         {
           \set{\beta}, x\geq 0,\subst{y\leftarrow x}}
          {s_2 \OTarrow {\alpha(x)} s_2'}
\qquad
\openrule
         {
           \set{\beta}, x<0,\subst{y\leftarrow x}}
          {s_2 \OTarrow {\alpha(x)} s_2'}
\]
This example illustrates the necessity of multiple transitions in the definition of bisimulation in a naive and minimalistic way.  It can easily be extended into a non-trivial example with more states and different usage of the variables.
\end{example}

@@

 $ \forall n\ge 1.  WI_{8b}(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(s\_msg,s\_ec}+n)\}, True, ()}
         {220 \OTWeakarrow {\nounderline{\texttt{out(s\_msg,m\_ec}+n)}} \{202, 000\}}
         $
